Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1) \note{\scriptsize Authors, please add/check middle initials and affiliations}

### 2014-04-13 WSSSPE1_report/WSSSPE1_summary_report.tex
Fine tune sections 3 and 4.

He concluded with a reference to his paper~\cite{bourne_ten} and
argued that computational scientists ``have a responsibility to
convince their institutions, reviewers, and communities that software is
scholarship, frequently
more valuable than a research article,'' a point with which all authors
strongly agree.

@@

In conclusion Smith reiterated the importance of establishing
incentive models for open contributions and tool builders, for
example, meaningful metrics and research grants such as~\cite{NSF_software_vision}. He urged computational scientists to
collaborate and share often their research reports, teaching
materials, code, as well as data by attaching proper licenses.

@@

The panel on developing and supporting software examined the challenges
around scientific software development and support, mainly focused on research
groups that also produce code in various forms. There was widespread agreement
that developing and maintaining software is hard, but best practices can help.
Several participants added that documentation is not just for users, and writing
API documentation, tutorials for building and deploying software, along with
documented development practices can be very helpful in bringing  new developers
into a project efficiently.

@@

Recruiting and/or retaining personnel in this area is hard,
with one of the major reasons being no long-term career paths (especially when
compared with industry). This is further discussed in \S\ref{sec:career-tracks}.

@@

The question of what ``sustainable'' means in the context of software was
raised (see \S\ref{sec:defining-sustainability}.) The resources at \url{http://oss-watch.ac.uk/resources/ssmm} discuss what
to look for when choosing software for procurement, or reuse in further software
development. Regardless of the license and development model that will be used,
the future of the project must be considered. Even if a particular piece of
software satisfies today's requirements, will it continue to do so in the
future? Important questions include whether the supplier will still be around in five
years time, will it still care for its customers needs, will it be responsive to
bug reports and feature requests? Should you tie your investment to a single
supplier using a proprietary product, or ensure the project uses an OSI-approved
license, and can outlive any single entity if the software is still important.

@@

The papers submitted to this panel, and several others, include lots of general
recommendations for processes, practices, tools, etc. One of the papers
suggested that a ``Software Sustainability Institute'' should be vested with
the authority to develop standardized quality processes, a central common
repository, central resources for services and consulting, a think tank of
sorts, and a software orphanage center. The idea of one common repository received
some resistance, with so many compelling alternatives available, e.g. Bitbucket or
GitHub. The point for centralization of communication/point of contact was seen as
reasonable, with the statement that ``vested with authority'' is perhaps too strong,
but ``providing tools if needed'' might be more appropriate.

>>>>

Bourne concluded with a reference to his paper~\cite{bourne_ten} and
argued that computational scientists ``have a responsibility to
convince their institutions, reviewers, and communities that software is
scholarship, frequently
more valuable than a research article,'' a point with which all authors
strongly agree.

@@

In conclusion Smith reiterated the importance of establishing effective
incentive models for open contributions and tool builders, for
example, meaningful metrics and research grants such as~\cite{NSF_software_vision}. He urged computational scientists to
collaborate and share often their research reports, teaching
materials, code, as well as data by attaching proper licenses.

@@

The panel on developing and supporting software examined the challenges
around scientific software development and support, mainly focused on research
groups that also produce code in various forms. There was widespread agreement
that developing and maintaining software is hard, but best practices can help.
Several participants added that documentation is not just for users, and writing
API documentation, tutorials for building and deploying software, together with
documented development practices can be very helpful in bringing  new developers
into a project efficiently.

@@

Recruiting and retaining personnel in this area is hard,
with one of the major reasons being no long-term career paths (especially when
compared with industry). This is further discussed in~\S\ref{sec:career-tracks}.

@@

The question of what ``sustainable'' means in the context of software was
raised (see~\S\ref{sec:defining-sustainability}). The resources at \url{http://oss-watch.ac.uk/resources/ssmm} discuss what
to look for when choosing software for procurement, or reuse in further software
development. Regardless of the license and development model that will be used,
the future of the project must be considered. Even if a particular piece of
software satisfies today's requirements, will it continue to do so in the
future? Important questions include whether the supplier will still be around in five
years' time, will it still care for its customers' needs, will it be responsive to
bug reports and feature requests? Should you tie your investment to a single
supplier using a proprietary product, or ensure the project uses an OSI-approved
license, and can outlive any single entity if the software is still important.

@@

The papers submitted to this panel, and several others, include lots of general
recommendations for processes, practices, tools, etc. One of the papers
suggested that a ``Software Sustainability Institute'' should be vested with
the authority to develop standardized quality processes, a central common
repository, central resources for services and consulting, a think tank of
sorts, and a software orphanage center. The idea of one common repository received
some resistance, with so many compelling alternatives available, e.g. Bitbucket or
GitHub. The centralized communication or point of contact was seen as
reasonable, with the statement that ``vested with authority'' is perhaps too strong,
but ``providing tools if needed'' might be more appropriate.
